null_model

### 2020-09-09 analysis/ts_null_model.Rmd
Null model in parallel.
Modify get_plot_sample and get_plot_FD for new data structure.

```{r}
library(tidyverse)
library(biodivTS)
library(FD)
```

@@

Simulation functions
```{r}
get_plot_sample <- function(data, filter_plot, n){
  plot_data <- data %>% filter(plot_id == filter_plot)

  rich_levels <- plot_data %>%
    select(year, id) %>%
    distinct() %>%
    count(year) %>%
    pull(n) %>%
    unique()

@@

  if (length(species_pool) < max(rich_levels)) browser()

  get_samp_species <- function(richness){
    samp_species <- sample(species_pool, richness)
    df <- cbind(richness, samp_species)
    return(df)
  }

  site_samples <- data.frame()
  for (j in 1:n){
    iter_samples <- data.frame()
    for(i in rich_levels){
    iter_samples <- rbind(iter_samples, get_samp_species(i))
    }
    site_samples <- rbind(site_samples, iter_samples %>% mutate(n = j))
  }
  site_samples %>%
    mutate(plot_id = filter_plot, value = 1) %>%
    pivot_wider(names_from = samp_species, values_from = value)
}

@@

Get dataframe of samples for each plot and corresponding species and trait matrices
```{r}
load(here::here("data", "biotime_traits.rda"))

@@

sample_occurrence <- map_dfr(unique(sample_plots$plot_id), get_plot_sample, data = sample_plots, n = n) %>%
  replace(is.na(.), 0)

#colnames(biotime_traits)[17:64]

get_plot_FD <- function(data, trait_range, occ_df, filter_plot) {
  plot_samples <- occ_df %>%
    filter(plot_id == filter_plot)

  sample_species_mat <- plot_samples %>%
    select(-c(richness, plot_id, n)) %>% #want only species x site info
    select_if(!colSums(.) %in% c(0, NA)) %>% #remove species that don't occur in the plot
    select(sort(tidyselect::peek_vars())) #sort columns to be in alphabetical order, required for dbFDI()

  fs::dir_create(paste0(here::here("data"), "/fd_output"))

  #the convexhull c code creates an output file of vertices, in order to not have multiple threads writing
  #to the same file we have to create individual working directories
  tmp_dir <- paste0(here::here("data", "fd_output", "tmp"), "/tmp_", filter_plot)
  fs::dir_create(tmp_dir)
  setwd(tmp_dir)

  print(filter_plot)
  sample_fd <- as.data.frame(dbFD_joggle(x = sample_trait_mat, a = sample_species_mat, w.abun = FALSE)) %>%
    cbind(plot_samples %>% select(richness, plot_id, n))

  save_dir <- paste0(here::here("data", "fd_output", "null"))
  fs::dir_create(save_dir)
  write_tsv(sample_fd, paste0(save_dir, "/null", filter_plot, ".csv"))

  df <- occ_df %>%
    filter(plot_id == filter_plot) %>%
    select(richness, plot_id, n) %>%
    bind_cols(sample_fd) %>%
    mutate(merge_test = case_when(richness == nbsp ~ TRUE,
                                  TRUE ~ FALSE))

  if (sum(df$merge_test) != dim(df)[1])
    warning("richness and region columns may not have been appropriately joined with FD data")

  return(df %>% select(-merge_test))
}

#test <- get_plot_FD(sample_plots, 18:65, sample_occurrence, 391)
null_FD <- map_dfr(unique(sample_plots$plot_id)[1:3], get_plot_FD, 
                   data = sample_plots, trait_range = 18:65, occ_df = sample_occurrence)
```

@@

traits <- euc.pco$points
row.names(traits) <- row.names(traits2)

```

>>>>

```{r}
library(tidyverse)
library(biodivTS)
library(FD)

pins::board_register_github(repo = "karinorman/biodivTS", branch = "master")
```

@@

Simulation functions
```{r}
get_plot_sample <- function(data, filter_plot, n){
  print(filter_plot)
  plot_data <- data %>% filter(plot_id == filter_plot) %>%
    select(year, id, year_abun)

@@

  get_samp_species <- function(data){
    samp_species <- sample(species_pool, n_distinct(data$id))
    samp_abun <- sample(data$year_abun)

    df <- as_tibble(cbind(samp_species, samp_abun)) %>% mutate(year = unique(data$year))
    return(df)
  }

  site_samples <- data.frame()
  for (j in 1:n){
    iter_samples <- data.frame()
    for(i in unique(plot_data$year)){
      year_data <- plot_data %>% filter(year == i)
      iter_samples <- rbind(iter_samples, get_samp_species(year_data))
    }
    site_samples <- rbind(site_samples, iter_samples %>% mutate(n = j))
  }
  site_samples %>%
    mutate(plot_id = filter_plot) %>%
    pivot_wider(names_from = samp_species, values_from = samp_abun)
}

@@

Get dataframe of samples for each plot and corresponding species and trait matrices
```{r}
biotime_traits <- pins::pin_get("biotime-traits", board = "github")

@@

# sample_occurrence <- map(unique(sample_plots$plot_id), get_plot_sample, data = sample_plots, n = n) %>%
#   bind_rows() %>%
#   replace(is.na(.), 0)
# pins::pin(sample_occurrence, board = "github")
sample_occurrence <- pins::pin_get("sample_occurrence", board = "github")

get_plot_FD <- function(data, occ_df, filter_plot, ...) {
  plot_samples <- occ_df %>%
    filter(plot_id == filter_plot)

  sample_species_mat <- plot_samples %>%
    select(-c(year, plot_id, n)) %>% #want only species x site info
    select_if(!colSums(na.omit(.)) %in% c(0, NA)) %>% #remove species that don't occur in the plot
    select(sort(tidyselect::peek_vars())) #sort columns to be in alphabetical order, required for dbFDI()

  trait_mat <- data %>%
    filter(id %in% colnames(sample_species_mat)) %>%
    select(-c(biotimeName, taxa, id_species, plot, year, year_abun, 
              year_biomass, traitName, scientificName, plot_id, study_id)
    ) %>%
    distinct() %>%
    select_if(~ length(unique(na.omit(.))) > 1)  #remove columns that have the same value for all columns

  #get binary variables so they can be excluded from rescaling 
  bin_vars <- map(trait_mat, ~ all(na.omit(.) %in% 0:1))

  trait_mat <- trait_mat %>%
    mutate_at(vars(-c(id, names(bin_vars[bin_vars == TRUE])), -ends_with("5cat")), 
              list(~as.numeric(scale(.)))) %>% #rescale variables, not binary or categorical
    arrange(id) %>%
    column_to_rownames("id")

  save_dir <- here::here("data", "fd_null_output")
  fs::dir_create(save_dir)

  #the convexhull c code creates an output file of vertices, in order to not have multiple threads writing
  #to the same file we have to create individual working directories
  tmp_dir <- paste0(save_dir, "/tmp", "/tmp_", filter_plot)
  fs::dir_create(tmp_dir)
  setwd(tmp_dir)

  print(filter_plot)
  sample_fd <- as.data.frame(dbFD_joggle(x = trait_mat, a = sample_species_mat, w.abun = TRUE, ...)) %>%
    cbind(plot_samples %>% select(year, plot_id, n))

  setwd(here::here())
  write_tsv(sample_fd, paste0(save_dir, "/null", filter_plot, ".csv"))

  # df <- occ_df %>%
  #   filter(plot_id == filter_plot) %>%
  #   select(richness, plot_id, n) %>%
  #   bind_cols(sample_fd) %>%
  #   mutate(merge_test = case_when(richness == nbsp ~ TRUE,
  #                                 TRUE ~ FALSE))
  # 
  # if (sum(df$merge_test) != dim(df)[1])
  #   warning("richness and region columns may not have been appropriately joined with FD data")

  return(sample_fd)
}

null_FD <- map_dfr(unique(sample_plots$plot_id)[1:3], get_plot_FD, 
                   data = sample_plots, occ_df = sample_occurrence,
                   m = "min", corr = "cailliez")
```

@@

traits <- euc.pco$points
row.names(traits) <- row.names(traits2)
```
